---
title: "a3_task3_shelby_smith.Rmd"
author: "Shelby Smith"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(kableExtra)
urchin_data<-read.csv(here("data","urchins.csv"))%>% #read in the urchin data
clean_names()
```
## 5.a Section 1: Size comparisons, purple and red urchins

### 5.a.i Exploratory data visualization:

```{r}
 ggplot(data = urchin_data,
        aes(x=size))+
  geom_histogram(bins = 15)+ #makes a histogram & shrinks x-axis to portion with values in it
  facet_wrap(~common_name)+ #splits into 2 graphs by common name
  labs(title = "Exploratory Histogram: Size Comparisons of Purple and Red Urchins",
       x= "Urchin Diameter (cm)",
       y= "Number of Urchins")

ggplot(data = urchin_data,
       aes(x=common_name, y=size))+
  geom_boxplot()+
  labs(x="Urchin Species Common Name",
       y= "Urchin Diameter (cm)",
       title = "Exploratory Boxplot: Size Companisons of Purple and Red Urchins")
```

### 5.a.ii Descriptive Statistics Table

```{r}
urchin_prelim_stats<- urchin_data %>% 
  group_by(common_name) %>% #tell r it's working with groups within the common_name column 
  summarize(mean_size = mean(size), #label the name of the new column = function you want to use (on what data)
            median_size = median(size),
            sd_size = sd(size),
            sample_size = n()) 
  
kbl(urchin_prelim_stats) %>% #use kableExtra package for good table viz in knitted doc
  kable_paper("hover", full_width=F)

#actual difference in mean urchin diameter between the species
actual_difference <- urchin_prelim_stats$mean_size[2] - urchin_prelim_stats$mean_size[1]

# Percent difference in mean urchin diameter between the species
percent_difference<- ((urchin_prelim_stats$mean_size[2]-urchin_prelim_stats$mean_size[1])/urchin_prelim_stats$mean_size[2])*100
```

### 5.a.iii Summary of Exploratory Data Visualizations and Tables

- The Red Urchin has a greater central tendency as compared to the more left skewed Purple Urchin histogram. 

- Red Urchins' size has a greater data spread, confirmed by the greater standard deviation in the table. 

- Contextually, although the mean diameter of the Red Urchin is 2.3 cm larger than the mean diameter of Purple Urchins, the Red Urchin is only 34% larger. 

## 5.b.i Section 2: Explore purple urchin counts by site and year

```{r}
purple_urchins <- urchin_data %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date=mdy(date)) %>%  
  #convert the date to date class "Date", actually replaces the original column 'date' with a new column 'date' but with the values in class 'Date'
  mutate(year=year(date),
         site_full= case_when(site == "NAPL" ~ "Naples Reef", 
                   site == "AQUE" ~"Arroyo Quemado Reef", 
                   site == "CARP" ~ "Carpinteria Reef", 
                   site == "MOHK" ~ "Mohawk Reef", 
                   site== "IVEE" ~ "Isla Vista Reef"))
#Create a new column called site_full that contains the full name (not abbreviation) for each of the five sites
```

### 5.b.ii Summary Table to total number of purple urchins observed by site and year

```{r}
summary_stats_purple_urchin<- purple_urchins %>% 
  group_by(year, site_full) %>% 
  count(common_name) %>% 
  rename(urchin_count=n)
```
### 5.b.iii Exploratory graph of counts of purple urchins recorded at each site over time for the years included in the study

```{r}
summary_stats_purple_urchin %>% 
  ggplot(aes(x=year, y=urchin_count, color=site_full)) + 
  geom_col(show.legend = FALSE) +
  facet_wrap(~site_full) +
  labs(x="Year",
       y= "Number of Urchins",
       title = "Purple Urchin per Year")+
  theme_minimal()
```



